%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%W  copyrigh.tex           SONATA documentation                     
%W                                                                 
%%
%Y  Copyright 2006,  Institut fuer Algebra, Universitaet Linz,  Austria
%%


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\Chapter{Copyright notice}

\begingroup
\def\"#1{\accent127 #1}
\parindent 0pt
\parskip 0.7ex

{\everypar{\hangindent\manindent}
Copyright {\copyright} 2006 by\hfill\break
Aichinger, E., Binder, F., Ecker, J., Mayr, P., and N{\"o}bauer, C.,\hfill\break 4040 Linz, Austria}

\medskip

 SONATA  is  distributed as a free package for GAP, Version 4;
 you can redistribute it and/or
 modify it under the terms of the GNU General Public License
 as published by the Free Software Foundation;
 either version 2 of the License, or (at your option) any later version.

 For details, see https://www.gnu.org/licenses/gpl.html.

 This program is distributed in the hope that it will be useful,
 but WITHOUT ANY WARRANTY; without even the implied warranty of
 MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
 GNU General Public License for more details.

 If you obtain SONATA please send us  a short notice to that effect,
 e.g., an  e-mail  message   to  the  address
 `sonata@algebra.uni-linz.ac.at',
 containing your full  name and address. 
 This  allows us to keep track of the number of SONATA users.

 If you  publish  a mathematical  result  that  was  partly obtained using
 SONATA, please cite SONATA, just as you would cite another paper that you
 used.\* We would appreciate if you could inform us  about such a paper.
 Also please let us know if you  modify any part of SONATA.


\vfill\hrule

\noindent\llap{\*\enspace}Specifically, please refer to

\smallskip \parindent4pc \parskip 0pt
\item{[SONATA]}
Aichinger, E., Binder, F., Ecker, J., Mayr, P., and N{\"o}bauer, C.,
\item{} SONATA --- system of near-rings and their applications, GAP-package, Version 2;
2003
\item{} Institut f\"ur Algebra, University of Linz, Austria.
\item{} (http://www.algebra.uni-linz.ac.at/Sonata/)
\endgroup

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%E  copyrigh.tex  . . . . . . . . . . . . . . . . . . . . . . . ends here

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End: 
