%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%W  copyrigh.tex              GAP documentation                      LDFM
%W                                                                 & SMCS
%%
%H  @(#)$Id: copyrigh.tex,v 1.3 2003/03/26 14:19:11 juergen Exp $
%%
%Y  Copyright 1997,  Lehrstuhl D fuer Mathematik,   RWTH Aachen,  Germany
%%


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\Chapter{Copyright notice}

\begingroup
\def\"#1{\accent127 #1}
\parindent 0pt
\parskip 0.7ex

{\everypar{\hangindent\manindent}
Copyright {\copyright} 1997 by\hfill\break
        Institut f\"ur Algebra, Stochastik und
        wissensbasierte mathematische Systeme,\hfill\break
        University of Linz, 4040 Linz, Austria}

\medskip

SONATA  is  distributed as a package for GAP, Version 4.  GAP  is a
system  for  computational  discrete algebra with  particular emphasis on
computational  group  theory.  As a GAP-package it  is  distributed
freely subject to GAP's 'copyleft' conditions, which  we thankfully adopt
and adapt for SONATA. 

SONATA   can be copied   and  distributed freely  for any  non-commercial
purpose.

If you copy SONATA for somebody else, you may ask this person for  refund
of your expenses.  This should cover cost of media, copying and shipping.
You are not allowed to ask for more than this.  In any case you must give
a copy of this copyright notice along with the program.

If you obtain SONATA please send us  a short notice to that effect, e.g.,
an  e-mail  message   to  the  address
`sonata-crew@bruckner.stoch.uni-linz.ac.at',
containing your full  name and address.  This  allows us to keep track of
the number of SONATA users.

If you  publish  a mathematical  result  that  was  partly obtained using
SONATA, please cite SONATA, just as you would cite another paper that you
used.\* Also   we would appreciate if you   could inform us  about such a
paper.

You  are permitted  to modify and  redistribute  SONATA,  but you are not
allowed  to restrict further redistribution.  That is to say  proprietary
modifications will  not  be allowed.  We want all  versions  of SONATA to
remain free.

If you  modify any part of SONATA and redistribute it,  you must supply a
`README'  document.   This should specify what modifications you made  in
which  files.  We do  not  want  to take  credit  or  be blamed  for your
modifications.

Of course we are interested in all of your modifications.  In  particular
we would like to see bug-fixes, improvements and new functions.  So again
we would appreciate it if you would inform us about all modifications you
make.

SONATA is distributed by us without any warranty, to the extent permitted
by applicable state law.  We  distribute SONATA *as is* without  warranty
of any kind, either expressed or implied, including,  but not limited to,
the implied  warranties  of merchantability and  fitness for a particular
purpose.

The entire risk as to the quality and performance of the program is  with
you.  Should SONATA prove defective, you assume the cost of all necessary
servicing, repair or correction.

In no  case  unless  required by applicable law will we, and/or any other
party who  may  modify  and  redistribute  SONATA  as permitted above, be
liable  to you for damages, including lost profits, lost monies or  other
special, incidental or consequential damages  arising out  of the  use or
inability to use SONATA.

\vfill\hrule

\noindent\llap{\*\enspace}Specifically, please refer to

\smallskip \parindent4pc \parskip 0pt
\item{[SON 97]}
        The SONATA Team, SONATA --- Systems Of Nearrings And Their
        Applications, Version 1, 1997
\item{} Institut f\"ur Algebra, Stochastik und wissensbasierte
        mathematische Systeme,\hfill\break
        University of Linz, Austria

\endgroup

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%E  copyrigh.tex  . . . . . . . . . . . . . . . . . . . . . . . ends here

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End: 
